---
title: "Bayesian Project"
author: "James Lee"
date: "12/5/2019"
output: word_document
---

# Bayesian Project

**Hierarchical Model**
The data of the project contains 5 factors:
Sex (2 levels)
Race (4 levels)
Age (5 levels)
Education (3 levels)
State (51 levels)
The total number of combinations of levels is 6120.
  

```{r}
dataPath <- 'C:/Users/jalee/Desktop/Bayesian/'
dat<-read.csv(paste(dataPath,"MScA_32014_BayesianMethods_CourseProjectData.csv",sep="/"))
str(dat)
```
  
**2 Model**
After running MCMC with the these data and the model obtain a Markov chain posterior sample for 870
parameters including 2-way interactions.
Each Markov chain of the stan object obama_fit has length 36000.

```{r}
load(paste(dataPath,"fit_ext_20160527_103144.Rdata",sep="/"))
library(shinystan)
```
  
**3 Analysis**
```{r}
library(HDInterval)
library(rstan)
MCMC<-rstan::extract(obama_fit)
sum.obama_fit<-rstan::summary(obama_fit)[[1]]
HDI95<-sum.obama_fit[,c(4,8)]
selection<-apply(HDI95,1,function(z) findInterval(0,z)!=1)
sum.obama_fit[selection,c(4,8)]

plot(obama_fit,pars=c("b_0"))
hist(MCMC$b_0)

plot(obama_fit,pars=c("b_sex"))
plot(obama_fit,pars=c("b_race"))
plot(obama_fit,pars=c("b_age"))
plot(obama_fit,pars=c("b_education"))
plot(obama_fit,pars=c("b_state"))
plot(obama_fit,pars=c("b_sex_race"))
```
  
## Questions
1. Find the most supportive groups corresponding to each of the main effects for
Barack Obama in 2012
2. Find the least supportive group corresponding to each of the main effects
3. Compare odds of approval by males grouped by race and education. Which of
such subgroups shows the highest and the lowest support for the candidate
4. Answer the same question, but for females  

**Main Effects**
Sex (2 levels)
Race (4 levels)
Age (5 levels)
Education (3 levels)
State (51 levels)  

```{r}
#Sex
sex<-MCMC$b_sex
colnames(sex)<-levels(dat$sex)
#Race
race<-MCMC$b_race
colnames(race)<-levels(dat$race)
#Age
age<-MCMC$b_age
colnames(age)<-levels(dat$age)
#Education
education<-MCMC$b_education
colnames(education)<-levels(dat$education)
#State
state<-MCMC$b_state
colnames(state)<-levels(dat$state)
#race_education (Q3,Q4)
race_education<-MCMC$b_race_education
dimnames(race_education)[[2]]<-levels(dat$race)
dimnames(race_education)[[3]]<-levels(dat$education)
#b_0
intercept<-MCMC$b_0
```

*Sex*
```{r}
sex_means<-apply(sex,2,function(z) mean(z+intercept))
sex_HDI<-apply(sex,2,function(z) hdi(z+intercept))
sexOdds<-rbind(mean=exp(sex_means),exp(sex_HDI))
sexOdds
```  
> Females were more supportive than males.  

*Race*
```{r}
race_means<-apply(race,2,function(z) mean(z+intercept))
race_HDI<-apply(race,2,function(z) hdi(z+intercept))
raceOdds<-rbind(mean=exp(race_means),exp(race_HDI))
raceOdds
which.max(raceOdds[1,])
```  
> Black population was the most supportive

*Age*
```{r}
age_means<-apply(age,2,function(z) mean(z+intercept))
age_HDI<-apply(age,2,function(z) hdi(z+intercept))
ageOdds<-rbind(mean=exp(age_means),exp(age_HDI))
ageOdds
which.max(ageOdds[1,])
which.min(ageOdds[1,])
```  
> 18-24 was the most supportive group.
> 55+ was the least supportive group.

*Education*
```{r}
education_means<-apply(education,2,function(z) mean(z+intercept))
education_HDI<-apply(education,2,function(z) hdi(z+intercept))
educationOdds<-rbind(mean=exp(education_means),exp(education_HDI))
educationOdds
which.max(educationOdds[1,])
which.min(educationOdds[1,])
```  
> College or More education level group was the most supportive group.
> No College education level group was the least supportive group.

*State*
```{r}
state_means<-apply(state,2,function(z) mean(z+intercept))
state_HDI<-apply(state,2,function(z) hdi(z+intercept))
stateOdds<-rbind(mean=exp(state_means),exp(state_HDI))
stateOdds
which.max(stateOdds[1,])
which.min(stateOdds[1,])
```
> District of Columbia was the most supportive state.
> Mississippi(MS) was the least supportive state.

*Male and Education Interaction*
```{r}
odds_male<-rep(intercept+sex[,1],12)
dim(odds_male)<-dim(race_education)
odds_male<-odds_male+race_education
odds_male[2,,]

#race main effects
race_matrix<-cbind(race,race,race)
dim(race_matrix)<-dim(race_education)
dimnames(race_matrix)<-dimnames(race_education)
race_matrix[2,,]

#education main effects
education_matrix<-cbind(education,education,education,education)
dim(education_matrix)<-dim(race_education)[c(1,3,2)]
education_matrix<-aperm(education_matrix,c(1,3,2))
dimnames(education_matrix)<-dimnames(race_education)
education_matrix[2,,]

#odds (final)
odds_male<-odds_male+race_matrix+education_matrix
odds_male[2,,]
(odds_male_mean<-apply(odds_male,c(2,3),mean))
(odds_male_hdi<-apply(odds_male,c(2,3),hdi))
exp(odds_male_mean)
exp(odds_male_hdi)
```  
>Among males, Black population with some college degree was the most supportive group.
>Among males, White population with no college degree was the least supportive group.

*Female and Education Interaction*
```{r}
odds_female<-rep(intercept+sex[,2],12)
dim(odds_female)<-dim(race_education)
odds_female<-odds_female+race_education+race_matrix+education_matrix
(odds_female_mean<-apply(odds_female,c(2,3),mean))
(odds_female_hdi<-apply(odds_female,c(2,3),hdi))
exp(odds_female_mean)
exp(odds_female_hdi)
```  
>Among females, black population with some college degree was the most supportive group.
>Among females, white population with no college degree was the least supportive group.







